Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-15 paper/paper.tex
pass

\begin{figure}[H]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-strings}Distribution of strings
    used as parameters of content options.}
\end{figure}

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, (1) vc. tem dados para dar
  suporte a isto (de http e tcp)? ou seja, quantas options do tipo
  ``content'' possuem modificadores.} \Luc{conseguirei esses dados}
For example, the creator of the rule may want to indicate that a
specific string be present in the header of an HTTP request
message. Content \emph{modifiers} serve this purpose. For example, the
modifier \CodeIn{http\_header} in the option \CodeIn{content:
  ``keep-alive''; http\_header} states the requirement that the string
``keep-alive'' be present in the header of the message. (A keep-alive
connection attribute indicates that a single TCP connection should
remain open for multiple HTTP requests/responses.)
Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
  depois do que falei acima...}

@@

\tname\ uses the frequency of content modifiers to discriminate more
or less likely rules. For example, historical data indicates that
content options that use the \CodeIn{http\_connection} modifier are
much less likely than content options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable for rules of any protocol in Suricatax,
currently, this heuristic function is only supported for tcp and http
in \tname. The rationale is that (1) these are the protocols involving
most attacks \Mar{Lucas, please confirm.} and (2) the payload of
messages of different protocols are not as segmented as the case of
tcp and http, which makes the reference to specific parts of the
payload less common for rule writers and more challenging for
\tname\Mar{Lucas, please confirm}.

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the contribution of
each rule. More precisely, we calculate this heuristic function with
the following formula:

\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{N}\]

\noindent
, where $N$ is the number of contents in a given rule and $M$ is the
number of content modifiers\Mar{isto eh confuso. vc. esta dizendo que
  o mesmo content pode ter mais de um modificador? ou a string poderia
  ser acessada por mais de um modificador. se for este o caso,
  sinceramente, achei isto bem confuso. afinal, a regra jah indica
  qual o modificador foi escolhido para acessar a string.}. \Luc{M eh
  o numero de modificadores associados a cada content. Temos
  content[1], content[2], ..., content[N]. Cada um desses contents tem
  modificadores associados content[i][1], content[i][2], ...,
  content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
  content. a frequencia seria 0, correto?}

\begin{figure}[H]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.\Mar{what about ``none'' (no modifier)?}}
\end{figure}

@@

\Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
que vai mante-la) e mostre como eh feito o calculo da regra desta
funcao heuristica para regra inteira.}

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''; http\_uri}, as the one from
Section~\ref{sec:jboss}. The contribution of this option to the score
of the corresponding rule would be $\frac{10,111}{10,111*1}$. The
value of $N$ for this case is 1 as the field of the HTTP message
\CodeIn{Request URI} holding the string can be accessed only through
the modifier \CodeIn{http\_uri}.

@@

\Mar{----------------------------------------------------------- estou
aqui}

A multiplicação por 1 no denominador eh referente a quantidade de
modificadores que estão associadas a esse content.\Mar{nao foi isto
  que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
  contents e o exemplo acima tem apenas 1 content.}  No caso da opção
content: "keep-alive", essa string pertence ao campo
"Connection:". Como falado acima, o suricata pode acessar esse campo
pelos modificadores http\_header ou http\_connection. Portanto, esse
content será pontuado de acordo com a popularidade desses dois
modificadores:
\[\frac{4089+29}{10111*2}\]

>>>>

\begin{figure}[t!]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-strings}Distribution of strings
    used as parameters of content options.}
\end{figure}

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{conseguirei esses dados} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)  Figure~\Fix{adicione pequena
  figura...verifique se ainda faz sentido depois do que falei
  acima...}

@@

\tname\ uses the frequency of content modifiers to discriminate more
or less likely rules. For example, historical data indicates that
content options that use the \CodeIn{http\_connection} modifier are
much less likely than content options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol in Suricata,
currently, \tname\ only supports tcp and http. These are the protocols
involving ~80\% of the attacks.

%% and
%% (2) the payload of messages of different protocols are not as
%% segmented as the case of tcp and http. Consequently, rule writers do
%% not refer to specific parts of the payload as often as the task o and
%% more challenging for

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the contribution of
each rule. More precisely, we compute this function with the following
formula:

\[\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]

\noindent
, where $N$ is the number of contents in a given rule and $M_i$ is the
number of content modifiers that could access the string associated
with content $i$. \Mar{Qual o rationale para considerar as varias
  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}%% Isto
                                                                             %% nao
                                                                             %% intuitivo
                                                                             %% para
                                                                             %% mim
                                                                             %% -Marcelo

%% \Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
%%   o mesmo content pode ter mais de um modificador? ou a string poderia
%%   ser acessada por mais de um modificador. se for este o caso,
%%   sinceramente, achei isto bem confuso. afinal, a regra jah indica
%%   qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado. Entretanto sim, uma "string poderia
%%   ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
%%   o numero de modificadores associados a cada content. Temos
%%   content[1], content[2], ..., content[N]. Cada um desses contents tem
%%   modificadores associados content[i][1], content[i][2], ...,
%%   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
%%   content. a frequencia seria 0, correto?} \Luc{Sim}

\begin{figure}[t!]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.\Mar{what about ``none'' (no modifier)?}}
\end{figure}

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''; http\_uri}, as the one from
Section~\ref{sec:jboss}. The contribution of this option to the score
of the corresponding rule would be $0.5=\frac{10,111}{10,111*1}$. The
value of $M_i$ for this case is 1 as the field of the HTTP message
\CodeIn{Request URI} holding the string can only be accessed through
the modifier \CodeIn{http\_uri}. For the option \CodeIn{content:
  ``keep-alive''; <mod>}, either \CodeIn{http\_header} or
\CodeIn{http\_connection} could substitute the placeholder
\CodeIn{<mod>} to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} in the header of an HTTP request
message\Mar{Lucas, a imagem mostrando header com campos Request URI e
  Connection e os conteudos citados acima pode ajudar.}. Consequently,
for this case, the constribution of content modifiers would be
$0.2=\frac{4089+29}{10111*2}$. Assuming that both options are
mentioned in a rule, the aggregated contribution of this heuristic
function would be \Fix{...}\Mar{Lucas, confirme que nao esta faltando
  calcular a media (dividir tudo por N).}

@@

%% \Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
%% que vai mante-la) e mostre como eh feito o calculo da regra desta
%% funcao heuristica para regra inteira.}\\
%% \Luc{-------------}

@@

%% A multiplicação por 1 no denominador eh referente a quantidade de
%% modificadores que estão associadas a esse content.\Luc{O correto eh ser dividido pela quantidade de modificadores. Atualizei a formula para tentar representar isso. Considere que no exemplo acima o content pudesse ser acessado por mais de um modificador. A pontuação dele então seria $\frac{10,111+x}{10,111*1}$ > 1. Para manter a pontuação entre 0-1, o certo nesse caso seria fazer $\frac{10,111+x}{10,111*2}$}\Mar{nao foi isto
%%   que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
%%   contents e o exemplo acima tem apenas 1 content.}
